---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
library(data.table)
library(lattice)
actO<-fread("activity.csv",colClasses=c("integer","POSIXct","integer"))
actO$date<-as.Date(actO$date,"%Y-%m-%d")

```
Original data looks like  

```{r}
str(actO)
head(actO)

```

## What is mean total number of steps taken per day?
```{r}
total_steps_each_day<-aggregate(steps~date,actO,sum,na.rm=T)
barplot(total_steps_each_day$steps,main="Total number of steps taken each day",xlab="Days",ylab="Total steps")
```

Mean of total steps is `r round(mean(total_steps_each_day$steps),2)`  
Median of total steps is `r median(total_steps_each_day$steps)`


## What is the average daily activity pattern?

```{r}
mean_steps_per_interval<-aggregate(steps~interval,actO,mean)
plot(mean_steps_per_interval$interval,mean_steps_per_interval$steps,type="l",main="Average number of steps taken",xlab="Interval",ylab="Mean steps")
max_steps_interval<-subset(mean_steps_per_interval,steps==max(steps))
abline(v=max_steps_interval$interval,col="red")

```

Interval `r max_steps_interval$interval` contains the maximum number of steps. It is also displayed on the above graph


## Imputing missing values

The total number of missing values in the original dataset is `r length(which(is.na(actO$steps)))`

```{r}
na_indices<-which(is.na(actO$steps))
mean_steps_each_day<-aggregate(steps~date,actO,mean,na.rm=T,na.action=na.pass)
actM<-actO
nnn<-sapply(na_indices,function(x){
  na_date<-actM[x,date]
  mean_step_for_date<-mean_steps_each_day$steps[mean_steps_each_day$date==na_date]
  if(is.nan(mean_step_for_date)) {
    mean_steps_each_day$steps[mean_steps_each_day$date==na_date]<<-0
  }
  actM[x,"steps"]<<-0
})
```

A new dataset that is equal to the original dataset but the missing values filled in by the mean value on that day is:  

```{r}
str(actM)
```
```{r}
total_steps_each_day2<-aggregate(steps~date,actM,sum)
barplot(total_steps_each_day2$steps,main="Total steps taken each day",xlab="Days",ylab="Total steps")
```

Mean of total steps per day is `r mean(total_steps_each_day2$steps)`  
Median of total steps per day is `r median(total_steps_each_day2$steps)`    
So, these values differ from values calculated above. Essentially, these values are smaller.  


## Are there differences in activity patterns between weekdays and weekends?

```{r}
actM$weektype<-apply(actM,1,function(x){
  wd<-weekdays(as.Date(x[[2]]))
  if(wd=="Sunday" | wd=="Saturday") {
    "weekend"
  }else {
    "weekday"
  }
  
})
actM$weektype<-factor(actM$weektype)

mean_steps_per_interval_per_weektype<-aggregate(steps~interval+weektype,actM,mean)
xyplot(steps~interval|weektype,data=mean_steps_per_interval_per_weektype,type="l",ylab="Number of steps",xlab="Interval")

```
